Data analysis for employee cluster descriptive statistics

### 2023-08-08 04. Analysis/scripts/03-03-data-analysis-employee-cluster-descriptive-statistics.Rmd
Update 03-03-data-analysis-employee-cluster-descriptive-statistics.Rmd

---
title: "Data analysis for employee cluster descriptive statistics"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    code_folding: hide
params:
  author: "Ben Jaques-Leslie"
  project_number: 2312
  project_name: "Occupancy Patterns"
  google_drive_home: "G:/Shared drives/MSG Projects/1.0 Real Estate Solutions (LQ1)/2312 Work Patterns"
  data_folder_1: "prepared-data"
  data_file_1: "clusters-employee-2023-06-08.rdata"
  data_file_2: "prep_employees_2023-05-24.rdata"
  in_number_of_clusters: 5
  in_specification: "scale_working_proportion_pairs_1_10"

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
tictoc::tic()
if(!("arsenal" %in% installed.packages()[,"Package"]))
{
  install.packages("arsenal")
}
if(!("geomtextpath" %in% installed.packages()[,"Package"]))
{
  install.packages("geomtextpath")
}
if(!("oesrrr" %in% installed.packages()[,"Package"]))
{
  devtools::install_github(repo = "GSA/oesrrr")
}
library(oesrrr)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(arsenal)
library(Hmisc)
library(readxl)
library(flextable)
library(ggthemes)
library(tidymodels)
```

@@

Filter to data with `r params$in_number_of_clusters` clusters with specification `r params$in_specification`.

```{r}
clust <- 
  kclusts %>% 
  filter(
    k == params$in_number_of_clusters &
      specification == params$in_specification
  )
```

@@

Filter to data with `r params$in_number_of_clusters` clusters with specification `r params$in_specification`.

```{r}
clusters <- 
  clust %>%
  unnest(cols = c(tidied))

assignments <- 
  clust %>% 
  unnest(cols = c(augmented))

clusterings <- 
  clust %>%
  unnest(cols = c(glanced))

@@

@@

```{r}
d <-
  assignments %>%
  select(k, person_id, .cluster, specification) %>%
  full_join(prep_employees) %>% 
  mutate(
    .cluster = fct_na_value_to_level(.cluster, level = "(Missing)")
  )
```
Review data.

@@

prep_a <- bind_rows(
  d %>%
    summaries() %>%
    mutate(group = "Whole cohort"),
  d %>%
    summaries(.cluster) %>%
    mutate(group = glue::glue("Cluster: {.cluster}")) %>%
    select(-.cluster),
  d %>% 
    summaries_rule_based(rule_based_all_remote),
  d %>% 
    summaries_rule_based(rule_based_primarily_teleworking_observation),
  d %>% 
    summaries_rule_based(rule_based_primarily_teleworking_working),
  d %>% 
    summaries_rule_based(rule_based_primarily_reporting_to_job_site_observation),
  d %>% 
    summaries_rule_based(rule_based_primarily_reporting_to_job_site_working),
  d %>% 
    summaries_rule_based(rule_based_mostly_teleworking_observation),
  d %>% 
    summaries_rule_based(rule_based_mostly_teleworking_working),
  d %>% 
    summaries_rule_based(rule_based_mostly_reporting_to_job_site_observation),
  d %>% 
    summaries_rule_based(rule_based_mostly_reporting_to_job_site_working),
  d %>% 
    summaries_rule_based(rule_based_position_category_missing),
  d %>% 
    summaries_rule_based(rule_based_position_category_onsite_flexible),
  d %>% 
    summaries_rule_based(rule_based_position_category_offsite),
  d %>% 
    summaries_rule_based(rule_based_position_category_onsite_required),
  d %>% 
    summaries_rule_based(rule_based_position_category_position_has_changed),
  d %>% 
    summaries_rule_based(rule_based_position_category_not_specified),
  d %>% 
    summaries_rule_based(rule_based_telework_onsite_minimum),
)
```

@@

```{r}
gg_density <- function(in_data,in_variable)
{
  in_variable_sym <- sym(in_variable)

  spec <- unique(in_data$specification)

  clus <- unique(in_data$k)

  in_data %>% 
  ggplot(aes(x = !!in_variable_sym, color = .cluster, fill = .cluster)) +
  geom_density(alpha = .25) +
    geomtextpath::geom_textdensity(aes(label = .cluster), 
                   hjust = "ymax", 
                   vjust = 1, text_only = TRUE, text_smoothing = 0, straight = TRUE) + 
    oescolorrrs::scale_color_oes(palette = 'accent colors') +
    oescolorrrs::scale_fill_oes(palette = 'accent colors') +
    theme_fivethirtyeight() +
    labs(
      title = glue::glue("Histogram of `{in_variable}`"),
      caption = glue::glue("For specification {spec} with {clus} clusters.")
    ) +
    theme(
      legend.position = "none",
      panel.background = element_rect(fill = "white"),
      plot.background = element_rect(fill = "white"),
      # panel.border = element_blank()
    )
}
```

@@

```{r}
num_vars <- 
  d %>% 
  select_if(is.numeric) %>% 
  select(-k) %>% 
  names()

@@

```{r}
num_vars[1:20] %>% 
  map(~gg_density(d,in_variable = .))
```

>>>>

---
title: "Data analysis for employee cluster descriptive statistics"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    code_folding: hide
params:
  author: "Ben Jaques-Leslie"
  project_number: 2312
  project_name: "Occupancy Patterns"
  google_drive_home: "G:/Shared drives/MSG Projects/1.0 Real Estate Solutions (LQ1)/2312 Work Patterns"
  data_folder_1: "prepared-data"
  data_file_1: "clusters-employee-2023-08-07.rdata"
  data_file_2: "prep_employees_2023-06-13.rdata"
  in_number_of_clusters_1: 2
  in_number_of_clusters_2: 5
  in_specification: "scale_observation_none_1_20_TRUE"

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
tictoc::tic()
if(!("arsenal" %in% installed.packages()[,"Package"]))
{
  install.packages("arsenal")
}
if(!("geomtextpath" %in% installed.packages()[,"Package"]))
{
  install.packages("geomtextpath")
}
if(!("oesrrr" %in% installed.packages()[,"Package"]))
{
  devtools::install_github(repo = "GSA/oesrrr")
}
if(!("oesrrr" %in% installed.packages()[,"Package"]))
{
  devtools::install_github(repo = "GSA/oesrrr")
}
library(oesrrr)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(arsenal)
library(Hmisc)
library(readxl)
library(flextable)
library(ggthemes)
library(tidymodels)
```

@@

Filter to data with `r params$in_number_of_clusters_1` clusters with specification `r params$in_specification`.

```{r}
clust_1 <- 
  kclusts %>% 
  filter(
    k == params$in_number_of_clusters_1 &
      specification == params$in_specification
  )
```

Filter to data with `r params$in_number_of_clusters_2` clusters with specification `r params$in_specification`.

```{r}
clust_2 <- 
  kclusts %>% 
  filter(
    k == params$in_number_of_clusters_2 &
      specification == params$in_specification
  )
```

@@

Filter to data with `r params$in_number_of_clusters_2` clusters with specification `r params$in_specification`.

```{r}
clusters_1 <- 
  clust_1 %>%
  unnest(cols = c(tidied))

assignments_1 <- 
  clust_1 %>% 
  unnest(cols = c(augmented))

clusterings_1 <- 
  clust_1 %>%
  unnest(cols = c(glanced))

@@

Check data.

```{r}
assignments_1 %>% 
  glimpse()
```

Filter to data with `r params$in_number_of_clusters_2` clusters with specification `r params$in_specification`.

```{r}
clusters_2 <- 
  clust_2 %>%
  unnest(cols = c(tidied))

assignments_2 <- 
  clust_2 %>% 
  unnest(cols = c(augmented))

clusterings_2 <- 
  clust_2 %>%
  unnest(cols = c(glanced))

Check data.

```{r}
assignments_2 %>% 
  glimpse()
```

Function to rename `.cluster` column to include the number of clusters.

```{r}
rename_to_k <- function(in_data)
{
  tot_cluster <- in_data %>% select(k) %>% distinct() %>% pull()
  out_data <-
    in_data %>% rename_with( ~ glue::glue("{.x}_{tot_cluster}"), .cluster)
  return(out_data)
}
```

Apply function to assignments and join.

```{r}
assignments <-
  assignments_1 %>%
  rename_to_k() %>%
  select(person_id, starts_with(".cluster")) %>%
  full_join(assignments_2 %>%
              rename_to_k() %>%
              select(person_id, starts_with(".cluster")))
```
Check data.

@@

Join to `prep_employees`

```{r}
d <-
  assignments %>%
  select(person_id, starts_with(".cluster")) %>%
  full_join(prep_employees) %>% 
  mutate(
    across(
      .cols = starts_with(".cluster"),
      .fns = ~fct_na_value_to_level(., level = "(Missing)")
    )
  )
```
Review data.

@@

prep_a <- bind_rows(
  d %>%
    summaries() %>%
    mutate(group = "Whole cohort"),
  d %>%
    summaries(.cluster_2) %>%
    mutate(group = glue::glue("Cluster 2: {.cluster_2}")) %>%
    select(-.cluster_2),
  d %>%
    summaries(.cluster_5) %>%
    mutate(group = glue::glue("Cluster 5: {.cluster_5}")) %>%
    select(-.cluster_5),
  d %>% 
    summaries_rule_based(rule_based_all_remote),
  d %>% 
    summaries_rule_based(rule_based_primarily_teleworking_observation),
  d %>% 
    summaries_rule_based(rule_based_primarily_teleworking_working),
  d %>% 
    summaries_rule_based(rule_based_primarily_reporting_to_job_site_observation),
  d %>% 
    summaries_rule_based(rule_based_primarily_reporting_to_job_site_working),
  d %>% 
    summaries_rule_based(rule_based_mostly_teleworking_observation),
  d %>% 
    summaries_rule_based(rule_based_mostly_teleworking_working),
  d %>% 
    summaries_rule_based(rule_based_mostly_reporting_to_job_site_observation),
  d %>% 
    summaries_rule_based(rule_based_mostly_reporting_to_job_site_working),
  d %>% 
    summaries_rule_based(rule_based_position_category_missing),
  d %>% 
    summaries_rule_based(rule_based_position_category_onsite_flexible),
  d %>% 
    summaries_rule_based(rule_based_position_category_offsite),
  d %>% 
    summaries_rule_based(rule_based_position_category_onsite_required),
  d %>% 
    summaries_rule_based(rule_based_position_category_position_has_changed),
  d %>% 
    summaries_rule_based(rule_based_position_category_not_specified),
  d %>% 
    summaries_rule_based(rule_based_telework_onsite_minimum),
)
```

@@

```{r}
gg_density <- function(in_data,in_variable,in_cluster_number)
{
  in_variable_sym <- sym(in_variable)

  # spec <- unique(in_data$specification)
  # 
  # clus <- unique(in_data$k)

  if(in_cluster_number==2)
  {
    step_01 <-
      in_data %>%
      ggplot(aes(
        x = !!in_variable_sym,
        color = .cluster_2,
        fill = .cluster_2
      )) +
      geom_density(alpha = .25) +
      geomtextpath::geom_textdensity(
        aes(label = .cluster_2),
        hjust = "ymax",
        vjust = 1,
        text_only = TRUE,
        text_smoothing = 0,
        straight = TRUE
      )
  }

  if(in_cluster_number==5)
  {
    step_01 <-
      in_data %>%
      ggplot(aes(
        x = !!in_variable_sym,
        color = .cluster_5,
        fill = .cluster_5
      )) +
      geom_density(alpha = .25) +
      geomtextpath::geom_textdensity(
        aes(label = .cluster_5),
        hjust = "ymax",
        vjust = 1,
        text_only = TRUE,
        text_smoothing = 0,
        straight = TRUE
      )
  }

  step_01 +
    theme_fivethirtyeight() +
    labs(
      title = glue::glue("Histogram of `{in_variable}`"),
      caption = glue::glue("For {in_cluster_number} clusters.")
    ) +
    theme(
      legend.position = "none",
      panel.background = element_rect(fill = "white"),
      plot.background = element_rect(fill = "white"),
      # panel.border = element_blank()
    )
}
```

@@

```{r}
num_vars <- 
  d %>% 
  select_if(is.numeric) %>% 
  names()

@@

```{r}
num_vars[1:20] %>% 
  map(~gg_density(d,in_variable = .,in_cluster_number = 2))
```

### Telework

#### Number of days

```{r}
gg_density(d,"daily_status_teleworking", in_cluster_number = 2)
```

```{r}
gg_density(d,"daily_status_teleworking", in_cluster_number = 5)
```

#### Proportion of observation window

```{r}
gg_density(d,"daily_status_teleworking_observation_prop", in_cluster_number = 2)
```

```{r}
gg_density(d,"daily_status_teleworking_observation_prop", in_cluster_number = 5)
```

### Reporting to job site

#### Number of days

```{r}
gg_density(d,"daily_status_reporting_to_job_site", in_cluster_number = 2)
```

```{r}
gg_density(d,"daily_status_reporting_to_job_site", in_cluster_number = 5)
```

#### Proportion of observation window

```{r}
gg_density(d,"daily_status_reporting_to_job_site_observation_prop", in_cluster_number = 2)
```

```{r}
gg_density(d,"daily_status_reporting_to_job_site_observation_prop", in_cluster_number = 5)
```

### Collocation with at least one team member

#### Number of days

```{r}
gg_density(d,"days_colocated_1", in_cluster_number = 2)
```

```{r}
gg_density(d,"days_colocated_1", in_cluster_number = 5)
```

#### Proportion of observation window

```{r}
gg_density(d,"days_colocated_1_observation_prop", in_cluster_number = 2)
```

```{r}
gg_density(d,"days_colocated_1_observation_prop", in_cluster_number = 5)
```

### Collocation with supervisor

#### Number of days

```{r}
gg_density(d,"days_colocated_supervisor", in_cluster_number = 2)
```

```{r}
gg_density(d,"days_colocated_supervisor", in_cluster_number = 5)
```

#### Proportion of observation window

```{r}
gg_density(d,"days_colocated_supervisor_observation_prop", in_cluster_number = 2)
```

```{r}
gg_density(d,"days_colocated_supervisor_observation_prop", in_cluster_number = 5)
```

### Collocation with teamn

#### Number of days

```{r}
gg_density(d,"days_colocated_team", in_cluster_number = 2)
```

```{r}
gg_density(d,"days_colocated_team", in_cluster_number = 5)
```

#### Proportion of observation window

```{r}
gg_density(d,"days_colocated_team_observation_prop", in_cluster_number = 2)
```

```{r}
gg_density(d,"days_colocated_team_observation_prop", in_cluster_number = 5)
```

### Whole team telewords

#### Number of days

```{r}
gg_density(d,"days_colocated_team", in_cluster_number = 2)
```

```{r}
gg_density(d,"days_colocated_team", in_cluster_number = 5)
```

#### Proportion of observation window

```{r}
gg_density(d,"days_colocated_team_observation_prop", in_cluster_number = 2)
```

```{r}
gg_density(d,"days_colocated_team_observation_prop", in_cluster_number = 5)
```

#### Number of days

```{r}
gg_density(d,"working_days", in_cluster_number = 2)
```

```{r}
gg_density(d,"working_days", in_cluster_number = 5)
```

#### Number of employees

```{r}
gg_density(d,"team_size_mode", in_cluster_number = 2)
```

```{r}
gg_density(d,"team_size_mode", in_cluster_number = 5)
```

#### Proportion of observation window

```{r}
gg_density(d,"daily_status_reporting_to_job_site_observation_prop", in_cluster_number = 2)
```

```{r}
gg_density(d,"daily_status_reporting_to_job_site_observation_prop", in_cluster_number = 5)
```

## Summary statistics

```{r}
d %>% 
  group_by(.cluster_2) %>% 
  skim(daily_status_reporting_to_job_site,
       daily_status_reporting_to_job_site_observation_prop,
       daily_status_teleworking,
       daily_status_teleworking_observation_prop,
       days_colocated_1,
       days_colocated_1_observation_prop,
       days_colocated_supervisor,
       days_colocated_supervisor_observation_prop,
       days_colocated_team,
       days_colocated_team_observation_prop,
       days_telework_team,
       days_telework_team_observation_prop,
       working_days,
       team_size_mode) %>% 
  as.data.frame()
```

```{r}
d %>% 
  group_by(.cluster_5) %>% 
  skim(daily_status_reporting_to_job_site,
       daily_status_reporting_to_job_site_observation_prop,
       daily_status_teleworking,
       daily_status_teleworking_observation_prop,
       days_colocated_1,
       days_colocated_1_observation_prop,
       days_colocated_supervisor,
       days_colocated_supervisor_observation_prop,
       days_colocated_team,
       days_colocated_team_observation_prop,
       days_telework_team,
       days_telework_team_observation_prop,
       working_days,
       team_size_mode) %>% 
  as.data.frame()
```
